---
title: "a3_task3_anastasia_kunz"
author: "Anastasia Kunz"
date: "10/31/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
library(tidyverse)
library(here)
library(janitor)
library(kableExtra)
library(lubridate)
```


# Task 3

```{r}
urchin_data <- read.csv(here("data", "urchins.csv")) %>% #read in data for this task
  clean_names()
```

### Urchin Size Comparisons 

##### Exploratory Data Visualizations
Compare distribution of sizes between purple and red urchins using faceted histograms and boxplots

Histogram:
```{r}
ggplot(urchin_data, aes(x = size)) + #plot urchin data and explore by size
  geom_histogram(bin = 20) + #first visualize by histogram
  facet_wrap(~common_name)+ # separate out by red and purple urchin, through common name
  labs(
    x = "Urchin Size (cm)",
    y = "Count",
    title = "Histogram of Urchin Size"
  ) + 
  theme_bw()
```

Box plot:

```{r}
ggplot(urchin_data, aes(x = common_name, y = size))+ # plot with common_name and size
  geom_boxplot() + #use geom box plot 
  labs(
    x = "Common Name",
    y = "Urchin Size (cm)",
    title = "Box Plot of Urchin Szie"
  )+ # label x and y axis
  theme_bw() #bw theme for clarity
```

##### Descriptive Statistics Table 

This summary table displays the mean, median, standard deviation and sample size for red and purple urchin samples. 

```{r}
urchin_stats <- urchin_data %>% #create new variable to easily call statistics for table below
  group_by(common_name) %>% #group data by the species common name so stats show up for purple and red urchins separately
  summarize(mean_size = mean(size), #use summarize to gather mean, med, std dev, and sample size of urchin samples
            median_size = median(size),
            standard_dev = sd(size),
            n = n())
urchin_stats %>%  
  kable(col.names = c("Species", #Use kable for clean table. Need species AND the statistics reported as col names.
                      "Mean (cm)",
                      "Median (cm)",
                      "Standard Deviation",
                      "Sample Size (n)")) %>% 
  kable_styling(full_width = FALSE)
```

#### Urchin Size Exploratory Data Analysis

In general, purple urchins seem to be smaller than red urchins as seen in both exploratory plots and the summary statistics table. According to the mean for the two samples, the Red Urchin has a diameter that is about 2.3 centimeters larger than the Purple Urchin. This is about 50% larger than the Purple Urchin. This is similar to the difference between the two species when comparing the sample medians. The Red Urchin median length is 2.5 cm greater in diameter than the Purple Urchin. Additionally, there is a greater spread of size for red urchins with a slightly higher standard deviation within the sample. Both populations look normal, with a slight skew in the purple urchin sample. 



### Purple Urchin Counts by Size and Year

```{r}
purple_urchins <- urchin_data %>% 
  filter(common_name == "Purple Urchin") %>% 
  mutate(date_converted = mdy(date)) %>% 
  mutate(year = year(date_converted)) %>% 
  mutate(site_full = case_when(
    site == "NAPL" ~ "Naples Reef",
    site == "AQUE" ~ "Arroyo Quemado Reef",
    site == "CARP" ~ "Carpinteria Reef",
    site == "MOHK" ~ "Mohawk Reef",
    site == "IVEE" ~ "Isla Vista Reef"
  ))

#Create summary table by urchin species showing year and site counts
purple_urchins %>% 
  group_by()
```




